Fully Automatic Loop-free Program Synthesis by Bounded Model Checking

### 2014-01-31 paper/paper.tex
Nudges
Make results table smaller

@@

Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.  Two example \newC programs are shown 
in Fig.~\ref{fig:c-}.

@@

@@

\begin{figure}[h]
{\small
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

@@

The exact details of how we encoding an $\mathcal{L}$-program are given in Sec.~\ref{sec:encode-l}.
We must now express the {\sc synth} and {\sc verif} formulae as safety properties
of \newC programs, which is given in Fig.~\ref{fig:c-synthverif}.

\begin{figure}[h]
\centering
\begin{minipage}[t]{.45\textwidth}
\begin{lstlisting}[language=C++]
void synth() {
  prog_t p = nondet();
  int in[N], out[M];

@@

The only wrinkle here is that we will sometimes synthesise a program containing
constants.  If we have synthesised a program for with $w=k$,
the constants in the program will be $k$-bits wide.  To extend the program
to an $n$-bit machine (with $n > k$), we need some way of deriving $n$-bit-wide
numbers from $k$-bit ones.  We have several strategies for this and
just try each in turn.  Our strategies are shown in Fig.~\ref{fig:generalize}.
$\mathcal{BV}(v, n)$ denotes an $n$-bit wide bitvector holding the value $v$
and $b \cdotp c$ means the concatenation of bitvectors $b$ and $c$.

@@

\paragraph{\bf {\sc Kalashnikov} synthesises a shorter program than {\sc Brahma}:}
This happens in 4 of the 29 cases.  This case is illustrated by benchmark p29, shown in Fig.~\ref{fig:obfuscated}.
The specification here is a piece of obfuscated code taken from the Conficker worm~\cite{conficker}, and
our goal is to synthesise an equivalent program which is easier to understand.  The obfuscated code
uses several tricks, including an apparently unbounded loop.  {\sc Brahma} is able to
produce an equivalent program consisting of four instructions using shifts and addition, whereas {\sc kalashnikov}
is able to produce the minimal program \verb|y * 45|.  It is also worth noting that the specification
fed to {\sc kalashnikov} was just the obfuscated code, with no further preprocessing needed.
\paragraph{\bf {\sc Kalashnikov} is unable to synthesise a program:}
This happens in 8 of the 29 cases.  In each case, {\sc Brahma} needs user guidance to synthesise the program.
The existence of instances that are beyond the reach of current automatic
methods, and thus require manual intervention, is to be expected.
% A total of 8 tests fall into this category.
\paragraph{\bf {\sc Brahma} and {\sc Kalashnikov} both synthesise minimal programs:}
In the remaining cases, both {\sc brahma} and {\sc kalashnikov} are able to produce
minimal programs.  We performed a Mann-Whitney U test (see Fig.~\ref{fig:stats-table})
to determine whether there was a significant difference
in the speed of the two tools.  The test was unable to reject the null hypothesis
(that the tools are equally fast) at the p=0.05
level, i.e. there is no statistically significant difference in the speed of
the two tools on this set of benchmarks.

@@

\iffalse
\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

 \caption{The {\sc 2Sum} algorithm for correctly rounded floating-point sums.}
  \label{fig:2sum}
\end{figure}
\fi

\begin{figure}
\begin{center}
{\small
\include{figures/results_table}
}
\end{center}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma} and~\cite{brahma-icse}.
 \label{fig:results-table}}
\end{figure}

@@

\begin{figure}
{\small
 \include{figures/stats_table}
}
 \caption{Performance statistics when {\sc Brahma} and {\sc Kalashnikov} were both successful.
 \label{fig:stats-table}}
\end{figure}

>>>>

\iffalse
\lstset{emph={
  assume, assert}, emphstyle=\bfseries}
\fi

@@

Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.

\iffalse
Two example \newC programs are shown 
in Fig.~\ref{fig:c-}.

@@

\end{figure}
\fi

@@

\begin{figure}[ht]
{\small
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

@@

The exact details of how we encode an $\mathcal{L}$-program are given in Sec.~\ref{sec:encode-l}.
We must now express the {\sc synth} and {\sc verif} formulae as safety properties
of \newC programs, which is given in Fig.~\ref{fig:c-synthverif}.

\begin{figure}[ht]
\centering
\begin{minipage}[t]{.45\textwidth}
\begin{lstlisting}[language=C++]
void synth() {
  prog_t p = nondet();
  int in[N], out[M];

@@

The only wrinkle here is that we will sometimes synthesise a program containing
constants.  If we have synthesised a program with $w=k$,
the constants in the program will be $k$-bits wide.  To extend the program
to an $n$-bit machine (with $n > k$), we need some way of deriving $n$-bit-wide
numbers from $k$-bit ones.  We have several strategies for this and
just try each in turn.  Our strategies are shown in Fig.~\ref{fig:generalize}.
$\mathcal{BV}(v, n)$ denotes an $n$-bit wide bitvector holding the value $v$
and $b \cdotp c$ means the concatenation of bitvectors $b$ and $c$.

@@

\paragraph{\bf {\sc Kalashnikov} synthesises a shorter program than {\sc Brahma}:}
This happens in 4 of the 29 cases, which is to be expected since {\sc Kalashnikov}'s outputs
are, by construction, guaranteed to be minimal.  This case is illustrated by benchmark p29, shown in Fig.~\ref{fig:obfuscated}.
The specification here is a piece of obfuscated code taken from the Conficker worm~\cite{conficker}, and
our goal is to synthesise an equivalent program which is easier to understand.  The obfuscated code
uses several tricks, including an apparently unbounded loop.  {\sc Brahma} is able to
produce an equivalent program consisting of four instructions using shifts and addition, whereas {\sc kalashnikov}
is able to produce the minimal program \verb|y * 45|.  It is also worth noting that the specification
fed to {\sc kalashnikov} was just the obfuscated code, with no further preprocessing needed.
\paragraph{\bf {\sc Kalashnikov} is unable to synthesise a program:}
This happens in 8 of the 29 cases.  In each case, {\sc Brahma} needs user guidance to synthesise the program.
The existence of instances that are beyond the reach of current automatic
methods, and thus require manual intervention, is to be expected.
% A total of 8 tests fall into this category.
\paragraph{\bf {\sc Brahma} and {\sc Kalashnikov} both synthesise minimal programs:}
In the remaining cases, both {\sc brahma} and {\sc kalashnikov} are able to produce
minimal programs.  It was not clear from just looking at the runtime numbers whether
any of the tools was significantly faster than the others, so we performed a Mann-Whitney U test.
Since the {\sc brahma} experiments were performed on a slower machine than the
{\sc kalashnikov} ones, we multiplied each of the runtimes by the CPU frequency of the machine
that was used for that experiment.  This gives an estimate of the time we would expect the test to
have taken on a 1GHz machine.  Scaling in this way  is more or less legitimate, since we have empirically
observed that our algorithm is CPU bound.

For each pair of tools ({\sc kalashnikov} vs. each of the {\sc brahma} configurations), the U test was
unable to reject the null hypothesis (that the tools are equally fast) at the p=0.05 level. In other words,
there is no statistically significant difference in the speed of {\sc kalashnikov} and {\sc brahma} on this
set of benchmarks, even after correcting for the difference in CPU frequency.

\begin{figure}[ht]

@@

\begin{figure}
\begin{center}
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc 2Sum
 \label{alg:2sum}}
\begin{algorithmic}
\Let{$s$}{$a+b$}
\Let{$b'$}{$s - a$}
\Let{$a'$}{$s - b'$}
\Let{$\delta b$}{$b - b'$}
\Let{$\delta a$}{$a - a'$}
\Let{$t$}{$\delta a + \delta b$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\linewidth}
\begin{algorithm}[H]
\caption{\sc Fast2Sum
 \label{alg:fast2sum}}
\begin{algorithmic}
\Require{$a \ge b$}
\Let{$s$}{$a + b$}
\Let{$z$}{$s - a$}
\Let{$t$}{$b - z$}
\Ensure{$a + b$ = $s + t$ exactly}
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{center}

 \caption{The {\sc 2Sum} algorithm for correctly rounded floating-point sums.}
  \label{fig:2sum}
\end{figure}

\begin{figure}
\centering
{\scriptsize
\include{figures/results_table}
}
\caption{Time for synthesis of machine integer benchmarks from~\cite{brahma} and~\cite{brahma-icse}.
 \label{fig:results-table}}
\end{figure}

@@
